---
title: "Assignment 1"
author: "Jeffrey Strickland"
date: "12/20/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Load Libraries
```{r}
library(dbplyr)
library(data.table)
```
#  run_analysis.R 
The run_analysis.R script performs the data preparation and then followed by the 5 steps required as described in the course project’s definition.

## Download the dataset
Dataset downloaded and extracted under the folder called `UCI HAR Dataset`

```{r}
filename <- "Coursera_DS3_Final.zip"
```

# Checking if archieve already exists.

```{r}
if (!file.exists(filename)){
  fileURL <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
  download.file(fileURL, filename, method="curl")
}  
```

# Checking if folder exists

```{r}
if (!file.exists("UCI HAR Dataset")) { 
  unzip(filename) 
}
```

## Assign each data to variables

```{r}
features <- read.table("UCI HAR Dataset/features.txt", col.names = c("n","functions"))
```

### The features selected for this database come from the accelerometer and gyroscope 3-axial raw signals tAcc-XYZ and tGyro-XYZ.

```{r}
activities <- read.table("UCI HAR Dataset/activity_labels.txt", col.names = c("code", "activity"))
```

### List of activities performed when the corresponding measurements were taken and its codes (labels)

```{r}
subject_test <- "test/subject_test.txt" # 2947 rows, 1 column
```

### contains test data of 9/30 volunteer test subjects being observed

```{r}
x_test <- read.table("UCI HAR Dataset/test/X_test.txt", col.names = features$functions)
```

### contains recorded features test data

```{r}
y_test <- read.table("UCI HAR Dataset/test/y_test.txt", col.names = "code")
```
### contains test data of activities’code labels

```{r}
subject_train <- read.table("UCI HAR Dataset/train/subject_train.txt", col.names = "subject")
```

### contains train data of 21/30 volunteer subjects being observed

```{r}
x_train <- read.table("UCI HAR Dataset/train/X_train.txt", col.names = features$functions)
```

### contains recorded features train data

```{r}
y_train <- read.table("UCI HAR Dataset/train/y_train.txt", col.names = "code")
```

### contains train data of activities’code labels

## Merges the training and the test sets to create one data set
The set X (10299 rows, 561 columns) is created by merging x_train and x_test using rbind() function

```{r}
X <- rbind(x_train, x_test)
```

The set Y (10299 rows, 1 column) is created by merging y_train and y_test using rbind() 

```{r}
Y <- rbind(y_train, y_test)
```

The set Subject (10299 rows, 1 column) is created by merging subject_train and subject_test using rbind() function

```{r}
Subject <- rbind(subject_train, subject_test)
```

The set Merged_Data (10299 rows, 563 column) is created by merging Subject, Y and X using cbind() function


```{r}
Merged_Data <- cbind(Subject, Y, X)
```
## Extracts only the measurements on the mean and standard deviation for each measurement
Set `TidyData` (10299 rows, 88 columns) is created by subsetting `Merged_Data`, selecting only columns: `subject`, `code` and the measurements on the `mean` and standard deviation (`std`) for each measurement

```{r}
TidyData <- Merged_Data %>% select(subject, code, contains("mean"), contains("std"))
```

## Uses descriptive activity names to name the activities in the data set
Entire numbers in `code` column of the `TidyData` replaced with corresponding activity taken from second column of the `activities` variable

```{r}
TidyData$code <- activities[TidyData$code, 2]
```

## Appropriately labels the data set with descriptive variable names
The `code` column in `TidyData` renamed into `activities`

```{r}
names(TidyData)[2] = "activity"
```

All `Acc` in column’s name replaced by `Accelerometer`

```{r}
names(TidyData)<-gsub("Acc", "Accelerometer", names(TidyData))
```
All `Gyro` in column’s name replaced by `Gyroscope`

```{r}
names(TidyData)<-gsub("Gyro", "Gyroscope", names(TidyData))
```

All `BodyBody` in column’s name replaced by `Body`

```{r}
names(TidyData)<-gsub("BodyBody", "Body", names(TidyData))
```
All `Mag` in column’s name replaced by `Magnitude`

```{r}
names(TidyData)<-gsub("Mag", "Magnitude", names(TidyData))
```

All start with character `f` in column’s name replaced by `Frequency`

```{r}
names(TidyData)<-gsub("^f", "Frequency", names(TidyData))
```

All start with character `t` in column’s name replaced by `Time`

```{r}
names(TidyData)<-gsub("^t", "Time", names(TidyData))
```

All `tbody` in column’s name replaced by `TimeBody`

```{r}
names(TidyData)<-gsub("tBody", "TimeBody", names(TidyData))
```

All `-mean()` in column’s name replaced by `Mean`

```{r}
names(TidyData)<-gsub("-mean()", "Mean", names(TidyData), ignore.case = TRUE)
```

All `-std()` in column’s name replaced by `Stand_Dev`

```{r}
names(TidyData)<-gsub("-std()", "STD", names(TidyData), ignore.case = TRUE)
```

## From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject

The set `FinalData` (180 rows, 88 columns) is created by sumarizing `TidyData` taking the means of each variable for each activity and each subject, after groupped by subject and activity.

```{r}
FinalData <- TidyData %>%
  group_by(subject, activity) %>%
  summarise_all(funs(mean))
```

Export `FinalData` into `FinalData.txt` file.

```{r}
write.table(FinalData, "FinalData.txt", row.name=FALSE)
```

### Print FianlData

```{r}
FinalData
```
